Research Notes

### 2018-11-02 writeups/notes.tex
check in new notes + higher k sims

Now, consider a hydraulic jump $u_1, h_1 \to u_2, h_2$ such that the
discontinuity is fixed (we've shifted to the jump front frame), so that
$\s*{uh}, \s*{hu^2 + gh^2/2}$ bracketed quantities are continuous across the jump.
This results in
\begin{align}
    h_1u_1^2 - h_2u_2^2 = h_1u_1^2\p*{1 - \frac{u_2}{u_1}}
        &= h_1u_1^2\p*{1 - \frac{h_1}{h_2}},\nonumber\\
        &= \frac{g\p*{h_2^2 - h_1^2}}{2},\nonumber\\
    u_1^2 &= \frac{h_2}{h_1}\frac{g\p*{h_1 + h_2}}{2}.
\end{align}
Customarily one defines $a_1^2 = gh_1, \mathrm{Fr} =
\frac{u_1}{a_1}$ and $\gamma = \frac{h_2}{h_1}$ which gives simplification
\begin{align}
    2\mathrm{Fr}^2 &= \gamma(1 + \gamma)\nonumber\\
    \gamma = \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
\end{align}
Note now that $\mathrm{Fr}^2 > 1$ produces a $\gamma > 1$, so $h_2 > h_1$.
Noting moreover that $\frac{u_1}{u_2} = \gamma$ and so $u_1 > u_2$. If we have a
picture of a jet incident on a medium at rest, $u_2$ is just the velocity of the
jump.

@@

This is the case worked out in detail in Bohr's paper. Following much the same
approach but in cylindrical coordinates we may write down equations of motion
\begin{align}
    u_r \pd{u_r}{r} + u_z \pd{u_r}{z} &= -g\pd{h}{r} + \nu
        \ptd{u_r}{z},\nonumber\\
    \frac{1}{r}\pd{(ru_r)}{r} + \pd{u_z}{z} &= 0.
\end{align}
Note the slight difference here due to the $\vec{\nabla} \cdot$ in cylindrical
coordinates. Interestingly, as a consequence of this $\pd{P_r}{t} +
\frac{1}{r}\partial_r \p*{hu_r^2 + gh^2/2} \neq 0$, since one term is a
$\vec{\nabla} \cdot (hu_r)$ and the other is a $\vec{\nabla}gh^2/2$! Radial
momentum is not conserved as the flow spreads out. Averaging using the same
techniques as above gives
\begin{align}
    \bar{u}_r\pd{\bar{u}_r}{r} + g\pd{h}{r} &= -\nu \frac{\bar{u}_r}{h^2},\nonumber\\
    \bar{u}_rhr &= q.
\end{align}
Combining the two gives
\begin{align}
    \pd{\bar{u}_r}{r}\p*{\bar{u}_r - \frac{gq}{\bar{u}_r^2r}}
        &= \frac{gq}{\bar{u}_rr^2} - \nu\frac{\bar{u}_r^3r^2}{q^2}.
        \label{eq:2d_hjump}
\end{align}
It's not very worth converting into dimensionless $\mathrm{Fr}$ number since it
would carry an $r$ dependence (if we are to hold $q$ constant) and we need to
study the $r$ dependence.

>>>>

Now, consider a hydraulic jump $u_1, h_1 \to u_2, h_2$ such that the
discontinuity is fixed (we've shifted to the jump front frame), so that
$\s*{uh}, \s*{hu^2 + gh^2/2}$ bracketed quantities are continuous across the jump.
This results in
\begin{align}
    h_1u_1^2 - h_2u_2^2 = h_1u_1^2\p*{1 - \frac{u_2}{u_1}}
        &= h_1u_1^2\p*{1 - \frac{h_1}{h_2}},\nonumber\\
        &= \frac{g\p*{h_2^2 - h_1^2}}{2},\nonumber\\
    u_1^2 &= \frac{h_2}{h_1}\frac{g\p*{h_1 + h_2}}{2}.
\end{align}
Customarily one defines $a_1^2 = gh_1, \mathrm{Fr} =
\frac{u_1}{a_1}$ and $\gamma = \frac{h_2}{h_1}$ which gives simplification
\begin{align}
    2\mathrm{Fr}^2 &= \gamma(1 + \gamma)\nonumber\\
    \gamma &= \frac{-1 + \sqrt{1 + 8\mathrm{Fr}^2}}{2}.
\end{align}
Note now that $\mathrm{Fr}^2 > 1$ produces a $\gamma > 1$, so $h_2 > h_1$.
Noting moreover that $\frac{u_1}{u_2} = \gamma$ and so $u_1 > u_2$. If we have a
picture of a jet incident on a medium at rest, $u_2$ is just the velocity of the
jump.

@@

This is the case worked out in detail in Bohr's paper. First, perhaps the
easiest thing to do is to enforce jump conditions $[rhu], [hu^2 + gh^2/2]$
subject to $q = 2\pi rh_iu_{ri}$ conserved total flow. This gives us
\begin{align}
    h_1u_{r1}^2 - h_2u_{r2}^2 = \frac{q}{2\pi r}(u_{r1} - u_{r2})
        &= \frac{gh_1^2}{2} - \frac{gh_2^2}{2},\\
        &= \frac{gq^2}{8\pi r^2u_{r1}^2} - \frac{gq^2}{8\pi r^2u_{r2}^2}
            ,\nonumber\\
    2u_{r1}^2 u_{r2}^2 &= \frac{gq}{2\pi r}\p*{u_{r2} + u_{r1}},\nonumber\\
    u_{r1} &= \frac{gq}{8\pi u_{r2}^2r} \p*{1 +
        \sqrt{1 + 16\frac{\pi u_{r2}^3r}{gq}}},\\
    \frac{u_{r1}}{u_{r2}} &= \frac{1}{4\mathrm{Fr}^2}\p*{1 +
        \sqrt{1 + 8\mathrm{Fr}^2}}.
\end{align}
We've defined $\mathrm{Fr}^2 \equiv \frac{2\pi r u_r^3}{gq} = \frac{u_r^2}{gh}$.
Note that this is in agreement with $\gamma = \frac{h_1}{h_2} = \frac{u_2}{u_1}$
from the plane parallel theory. The major difference is that now $\mathrm{Fr}^2
= \mathrm{Fr}^2(r)$, and so the location of the hydraulic jump can be pinned.
One might expect that once $\mathrm{Fr}(r) > 1$ that the jump occurs, but this
turns out not to be the case.

We now perform the boundary layer analysis. Following much the same approach but
in cylindrical coordinates we may write down equations of motion
\begin{align}
    u_r \pd{u_r}{r} + u_z \pd{u_r}{z} &= -g\pd{h}{r} + \nu
        \ptd{u_r}{z},\nonumber\\
    \frac{1}{r}\pd{(ru_r)}{r} + \pd{u_z}{z} &= 0.
\end{align}
Note the slight difference here due to the $\vec{\nabla} \cdot$ in cylindrical
coordinates. Interestingly, as a consequence of this $\pd{P_r}{t} +
\frac{1}{r}\partial_r \p*{hu_r^2 + gh^2/2} \neq 0$, since one term is a
$\vec{\nabla} \cdot (hu_r)$ and the other is a $\vec{\nabla}gh^2/2$! Radial
momentum is not conserved as the flow spreads out. Averaging using the same
techniques as above gives
\begin{align}
    \bar{u}_r\pd{\bar{u}_r}{r} + g\pd{h}{r} &= -\nu \frac{\bar{u}_r}{h^2},\nonumber\\
    \bar{u}_rhr &= q.
\end{align}
Combining the two gives
\begin{align}
    \pd{\bar{u}_r}{r}\p*{\bar{u}_r - \frac{gq}{\bar{u}_r^2r}}
        &= \frac{gq}{\bar{u}_rr^2} - \nu\frac{\bar{u}_r^3r^2}{q^2}.
        \label{eq:2d_hjump}
\end{align}
It's not very worth converting into dimensionless $\mathrm{Fr}$ number since it
would carry an $r$ dependence (if we are to hold $q$ constant) and we need to
study the $r$ dependence.
